\plaintitle

### 2017-04-28 statistical-quorum.tex
Small improvements.

% Paper metadata (use plain text, for PDF inclusion and later
% re-using, if desired).  Use \emtpyauthor when submitting for review
% so you remain anonymous.
\def\plaintitle{Using statistical quorum to improve legitimacy of decisions and utilization of collective attention in democratic communities}
%\def\plainauthor{First Author, Second Author, Third Author,
%  Fourth Author, Fifth Author, Sixth Author}
%\def\emptyauthor{}
%\def\plainkeywords{Authors' choice; of terms; separated; by
%  semicolons; include commas, within terms only; required.}
%\def\plaingeneralterms{Documentation, Standardization}

@@

@@

Many democratic communities use voting to make collective decisions.
\todo{Should we support this statement with a reference?}
In its basic form voting consist of members of a community casting a vote \emph{in favor} or \emph{against} a decision and if
majority ($> 50\%$) of votes are in favor of the decision, the decision \emph{passes} and is adopted as a decision
of the whole community.

There have been many improvements proposed to this basic form and various analysis made.
\todo{Add references.}
In this paper we focus on the question of \emph{quorum} and how high the quorum should be.
Quorum is the minimum number of members participating in voting for voting to be seen as representative of the whole
community, i.e., legitimate.
If quorum is too low, decisions adopted might not represent the community.
If it is too high, the community might have issues achieving the quorum.
Not enough members might decide to participate in voting, or members might on purpose boycott voting to
prevent decision to be adopted, even if those opposing it are in minority.
Communities often decide quorum arbitrary\footnote{Robert's Rules of Order states that the quorum
``should approximate the largest number that can be depended on to attend any meeting except in very bad weather
or other extremely unfavorable conditions.''~\cite{roberts}}
and try to increase participation of members in communities' democratic processes through various means
to address the issue of too low turnout.
\todo{Add references. Mention forced participation in some countries.}
Moreover, many democratic communities have issues achieving quorum which hinders their operation.
\todo{Add references.}

@@

To answer this question we are proposing a novel metric, \emph{decision credibility}, which uses statistical
hypothesis testing to evaluate the credibility of the voting result.
It uses distribution of votes made to evaluate the probability of decision passing if everyone actually voted,
and not just those who voted.
Intuitively, if in a community of 100 members 40 members vote for a decision to pass and nobody against,
we can see that if everyone voted, the probability of majority of population voting for a decision is high.
On the other hand, if 41 members for in favor and 40 against, we cannot really say much about what would a decision
be if everyone voted.
Only two more votes against can completely change the outcome.
Hence, the suggested action would be for the community to gather more votes and continue discussion around the decision.
We describe decision credibility in more detail in \secref{sec:statistical-quorum}.

>>>>

% Paper metadata (use plain text, for PDF inclusion and later
% re-using, if desired).  Use \emtpyauthor when submitting for review
% so you remain anonymous.
\def\plaintitle{Using decision credibility to improve legitimacy of decisions and utilization of collective attention in democratic communities}
%\def\plainauthor{First Author, Second Author, Third Author,
%  Fourth Author, Fifth Author, Sixth Author}
%\def\emptyauthor{}
%\def\plainkeywords{Authors' choice; of terms; separated; by
%  semicolons; include commas, within terms only; required.}
%\def\plaingeneralterms{Documentation, Standardization}

@@

%%% GLOSSARY %%%
%
% community = group which votes
% proposal = what you vote on
% decision = result of a vote, accepted or rejected proposal
%
%%% END GLOSSARY %%%

@@

Many communities use voting to make collective decisions.
\todo{Should we support this statement with a reference?}
In its basic form, voting consists of members of a community casting a vote \emph{in favor} or \emph{against} a decision.
If the majority ($> 50\%$) of votes are in favor of the decision, the decision \emph{passes} and is adopted as a decision
of the whole community.

There have been many improvements proposed to this basic form and various analysis made.
\todo{Add references.}
In this paper we focus on the question of \emph{quorum} and how high the quorum should be.
Quorum is the minimum number of members participating in voting for voting to be seen as representative of the whole
community, i.e., legitimate.
If quorum is too low, decisions adopted might not represent the community.
If it is too high, the community might have issues achieving the quorum.
Not enough members might decide to participate in voting, or members might on purpose boycott voting to
prevent decision to be adopted, even if those opposing it are in minority.
Communities often decide quorum arbitrarily\footnote{Robert's Rules of Order states that the quorum
``should approximate the largest number that can be depended on to attend any meeting except in very bad weather
or other extremely unfavorable conditions.''~\cite{roberts}}
and try to increase participation of members in communities' democratic processes through various means
to address the issue of too low turnout.
\todo{Add references. Mention forced participation in some countries.}
Moreover, many democratic communities have issues achieving quorum which hinders their operation.
\todo{Add references.}

@@

To answer this question we are proposing a novel metric, \emph{decision credibility}, which uses statistical
hypothesis testing to evaluate the credibility of the voting result.
It uses the distribution of votes made to evaluate the probability of a decision passing if everyone actually voted,
and not just those who did vote.
Intuitively, if, in a community of 100 members, 40 members vote for a decision to pass and nobody against,
we can see that if everyone voted, the probability of majority of population voting for a decision is high.
On the other hand, if 41 members for in favor and 40 against, we cannot really say much about what would a decision
be if everyone voted.
Only two more votes against can completely change the outcome.
Hence, the suggested action would be for the community to gather more votes and continue discussion around the decision.
We describe decision credibility in more detail in \secref{sec:statistical-quorum}.
